Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-30 manuscripts/reply/reply-main.tex
More work on intro for reply ms.

Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
of vertebrate lineages distributed across the Philippines; they also studied
the behavior of the \msb approach using computer simulations.
They found the method was very sensitive to prior assumptions and often
incorrectly supported shared divergences across taxa that diverged randomly
over broad time periods, to which \citet{Hickerson2013} responded.
\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
methodological point about the model selection performed in \msb:
% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
% on the fundamental methodological point about the model selection performed in \msb:
\begin{itemize}
   \item The use of vague priors on the divergence-time parameters can lead to
       spurious support for models with few divergence events shared across
       taxa. Thus, the primary inference enabled by the approach is very
       sensitive to the priors on divergence times.
\end{itemize}
However, the two papers suggest alternative mechanisms by which the priors on
divergence times cause this behavior:
\begin{itemize}
    \item \citet{Hickerson2013} suggest that under broad uniform priors, the
        rejection algorithm implemented in \msb is unable to adequately sample
        the space of the models within reasonable computational time.
        They propose the method incorrectly supports models with fewer
        divergence-time parameters because they are better sampled.
    \item \citet{Oaks2012} suggest the cause is more fundamental; the broad
        uniform priors simply lead to very small marginal likelihoods of models
        with many divergence-time parameters.
        Models with more divergence-time parameters integrate over \emph{much}
        greater parameter space with low probability of producing the data, yet
        relatively high prior density, resulting in small marginal likelihoods.
\end{itemize}
In the former case, the true (correct) posterior supports models with many
divergence-time parameters given data generated under temporally random
divergences, but the problem arises from our inability to accurately estimate
the posterior.
In the latter case, estimation error is not the issue, because the true
posterior supports simultaneous divergences given data from randomly diverged
taxa; this is the correct (albeit biogeographically misleading) answer given
the prior assumptions about divergence times imposed by broad uniform priors.
Such posterior support for simultaneous divergence, even if correct from the
perspective of Bayesian model choice, does not provide the biogeographical
insights that a researcher who employs \msb seeks to gain.

>>>>

Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
of vertebrate lineages distributed across the Philippines; they also studied
the behavior of the \msb approach using computer simulations.
They found the method was very sensitive to prior assumptions and often
incorrectly supported shared divergences across taxa that diverged randomly
over broad time periods, to which \citet{Hickerson2013} responded.
\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
methodological point about the model selection performed in \msb:
% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
% on the fundamental methodological point about the model selection performed in \msb:
\begin{itemize}
   \item The use of vague priors on the divergence-time parameters can lead to
       spurious support for models with few divergence events shared across
       taxa. Thus, the primary inference enabled by the approach is very
       sensitive to the priors on divergence times.
\end{itemize}
However, the two papers suggest alternative mechanisms by which the priors on
divergence times cause this behavior:
\begin{itemize}
    \item \citet{Hickerson2013} suggest that under broad uniform priors, the
        rejection algorithm implemented in \msb is unable to adequately sample
        the space of the models within reasonable computational time.
        They propose the method incorrectly supports models with fewer
        divergence-time parameters because they are better sampled.
    \item \citet{Oaks2012} suggest the cause is more fundamental; the broad
        uniform priors simply lead to very small marginal likelihoods of models
        with many divergence-time parameters.
        Models with more divergence-time parameters integrate over \emph{much}
        greater parameter space with low probability of producing the data, yet
        relatively high prior density, resulting in small marginal likelihoods.
\end{itemize}
In the former case, the problem is estimation error; the true (correct)
posterior supports models with many divergence-time parameters given data
generated under temporally random divergences, but we are unable to accurately
estimate the posterior.
In the latter case, estimation error is not the issue, because the true
posterior supports simultaneous divergences given data from randomly diverged
taxa; this is the correct (albeit biogeographically misleading) answer given
the prior assumptions about divergence times imposed by broad uniform priors.
Such posterior support for simultaneous divergence, even if correct from the
perspective of Bayesian model choice, does not provide the biogeographical
insights that a researcher who employs \msb seeks to gain.
